---
title: "OSC Herbarium"
author: "James Mickley"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(knitr) # Needed to print tables with kable()
library(taxize) # Needed to derive higher taxonomic ranks with tax_name()
library(sf) # Needed to work with shapefiles

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
data <- read.csv("data/osc-vascularplants.csv")

# Import Oregon Counties shapefile
or.counties <- read_sf("data/shapefiles/county.shp")


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size, move away from plot, and center
    plot.title = element_text(size = 22, face = "bold", vjust = 4, hjust = 0.5),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank(), 
    
    # Facet label text size and background
    strip.text = element_text(size = 16, face = "bold"),
    strip.background = element_rect(fill = NA)
    )


# Theme for maps to minimize clutter
map.theme <- theme(
    
    # Turn off axes
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    
    # Turn off gridlines, etc.
    #panel.border = element_blank(),
    #panel.grid.major = element_line(colour = "transparent"), 
    #panel.grid.minor = element_line(colour = "transparent"), 
    #panel.background = element_rect(fill = "transparent", colour = NA)
    )


```


```{r Data_Wrangling, echo = F, comment = ""}

# Get higher taxon information from ITIS
families <- data %>% 
    
    # Get a vector of unique family names
    pull(Family) %>% as.character() %>% unique() %>%
    
    # Get the order and division from ITIS
    tax_name(get = c("division", "class", "order"), db = "itis", 
        messages = F) %>% 
    
    # Rename columns
    select(Family = query, Division = division, Class = class, Order = order)


# Clean up the main dataset
data <- data %>%
    
    # Select columns to keep
    select(OccurrenceID, Herbarium, Collection, Catalog, Family, Genus, 
        ScientificName, Genus, SpecificEpithet, YearCollected, County, 
        DecimalLatitude, DecimalLongitude, CoordinateUncertaintyInMeters) %>% 
    
    # Add a column for decade
    mutate(Decade = floor(YearCollected/10)*10) %>%
    
    # Add a column for species without infraspecific info
    mutate(Species = paste(Genus, SpecificEpithet)) %>%

    # Add higher taxon information
    left_join(families, by = "Family") %>% 
    
    # Filter to vascular plants
    filter(Division == "Tracheophyta") %>%
    
    # Group by species to add some summaries
    group_by(Species) %>%
    
    # Add columns
    mutate(
        
        # Add the number of specimens for each species
        Specimens = n(), 
        
        # Add the last year each species was collected
        LastCollected = first(YearCollected, order_by = desc(YearCollected))
        )


# Clean up the oregon counties shapefile
or.counties <- or.counties %>% 
    
    # Rename NAME column
    select(County = NAME) %>% 
    
    # Transform county names from all caps
    mutate(County = str_to_title(County)) %>% 
    
    # Convert shapefile to wgs84
    st_transform(4326)


# Summarize county data
county.counts <- data %>% 
    
    # Group by species and county
    group_by(Species, County) %>%
    
    # Summarize the number of specimens of each species by county
    summarize(Specimens = n()) %>%
    
    # Group[ by county]
    group_by(County) %>%
    
    # Count the number of species and specimens in each county
    summarize(TotalSpp = n(), Specimens = sum(Specimens)) %>% 
    
    # Arrange counties by number of specimens (most to least)
    arrange(desc(Specimens))

```

## Summaries

```{r State_Summary, echo = F, comment = ""}

# Summarize total families, species, and specimens in dataset
data %>% 
    
    # Group by species
    group_by(Family, Species) %>%
    
    # Count up the number of specimens for each species
    summarize(Specimens = n()) %>%

    # Group by family
    group_by(Family) %>%
        
    # Summarize the # of species and specimens by family
    summarize(Species = n(), Specimens = sum(Specimens)) %>%
        
    # Summarize the # of families
    summarize(Families = n(), Species = sum(Species), 
        Specimens = sum(Specimens)) %>%

    # Output a table of the data
    kable(format = "markdown")  


# Summarize for common families (> 2000 specimens)
data %>% 
    
    # Group by species
    group_by(Family, Species) %>%
    
    # Count up the number of specimens for each species
    summarize(Specimens = n()) %>%

    # Group by family
    group_by(Family) %>%
        
    # Summarize the # of species and specimens by family
    summarize(Species = n(), Specimens = sum(Specimens)) %>%
    
    arrange(desc(Specimens)) %>%
    
    filter(Specimens > 2000) %>%

    # Output a table of the data
    kable(format = "markdown")  




data %>%
    ggplot() + 
        geom_sf(data = or.counties, color = "black") + 
        geom_point(aes(x = DecimalLongitude, y = DecimalLatitude), 
            color = "darkgreen", size = 2, alpha = 0.3) +
        theme_cowplot() +
        ggplot.theme +
        map.theme


```


### Specimens and species by county



## Sampling Biases

### Spatial

```{r Specimens_County, echo = F}



county.counts %>%
    ggplot(aes(x = reorder(County, -Specimens), y = Specimens)) + 
        geom_bar(stat = "identity") + 
        xlab("County") +
        theme_cowplot() + 
        ggplot.theme + 
        theme(axis.text.x = element_text(angle = 90, vjust = 0.2))


or.counties %>% left_join(county.counts) %>%
  ggplot() + 
    
    # Add the outline of CT, with counties
    geom_sf(aes(fill = Specimens), color = "black") + 
    geom_sf_label(aes(label = County), size = 4, fontface = "bold") + 
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +

    theme_cowplot() +
    ggplot.theme +
    map.theme

        
```

Species diversity at the county level looks partly tied to the number of specimens. Counties with more specimens have more diversity. This suggests that sampling is incomplete at the county level. 

```{r Diversity_County, echo = F}

or.counties %>%
    left_join(county.counts) %>%
    ggplot() + 
        
        geom_sf(aes(fill = TotalSpp), color = "black") + 
    
        geom_sf_label(aes(label = County), size = 4, fontface = "bold") + 
        scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
        theme_cowplot() +
        ggplot.theme +
        map.theme

```

### Temporal

```{r Specimens_Year, echo = F}

data %>% 
    group_by(YearCollected) %>% 
    summarize(Specimens = n()) %>%
    ggplot(aes(x = YearCollected, y = Specimens)) + 
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        theme_cowplot() + 
        ggplot.theme


```

```{r Specimens_Decade, echo = F}

data %>% 
    group_by(Decade) %>% 
    summarize(Specimens = n()) %>%
    ggplot(aes(x = Decade, y = Specimens)) + 
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        theme_cowplot() + 
        ggplot.theme


```

### Spatiotemporal

```{r Specimens_County_Decade, echo = F, fig.width = 12, fig.height = 9}

data %>% 
    group_by(County, Decade) %>% 
    summarize(Specimens = n()) %>%
    ggplot(aes(x = Decade, y = Specimens)) + 
        facet_wrap(~ County) +
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        theme_cowplot() + 
        ggplot.theme




```

```{r Specimens_County_Decade_Subset, echo = F}

data %>% filter(County %in% c("Lane", "Benton", "Harney",
    "Linn", "Columbia", "Baker")) %>% 
    group_by(County) %>%
    mutate(Total = n()) %>%
    group_by(County, Total, Decade) %>% 
    summarize(Specimens = n()) %>%
    ggplot(aes(x = Decade, y = Specimens)) + 
        facet_wrap(~ reorder(County, -Total)) +
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        theme_cowplot() + 
        ggplot.theme

```

### Undersampled Species

Undersampled Species
One way to get at data-deficient species, is to look at the number of species with very few records.

Of the 4500 species recorded, over 2000 less than 10 specimens statewide, and nearly 1000 species are only represented by 1-2 specimens. These would be a good target, though it's possible some are outdated names or something of that sort.

Notably though, most of the undersampled species were collected recently. There are pockets of undersampled species that haven't been collected recently in the CT River Valley, New London, and the NW corner that might be worth looking into.

```{r Undersampled_Species, echo = F}

# Filter to only include species with 10 or fewer specimens
data %>% 
    group_by(Species) %>%
    
    summarize(Specimens = n()) %>%
    
    filter(Specimens <= 10) %>%
    
    # Group by the number of specimens
    group_by(Specimens) %>% 
    
    # Summarize the number of species with a certain number of specimens
    summarize(Species = n()) %>% 
    
    # Output a table of the data
    kable(format = "markdown")




data %>% 
    
    group_by(Species) %>%
    
    summarize(Specimens = n()) %>%
    
    # Group by the number of specimens
    group_by(Specimens) %>% 
    
    # Summarize the number of species with a certain number of specimens
    summarize(Species = n()) %>% 
    
    mutate(MeanSpp = median(Species)) %>%
    
    ggplot(aes(x = Specimens, y = Species)) + 
    
        geom_bar(stat = "identity") + 
    
        geom_vline(aes(xintercept = MeanSpp), color = "red", size = 1.5) +
    
    theme_cowplot() + 
    ggplot.theme


```


```{r Undersampled_Species_LastCollected, echo = F}

data %>% 
    group_by(Species, LastCollected, Specimens) %>%
    summarize() %>% 
    filter(Specimens <= 10) %>%
    
    ggplot(aes(x = LastCollected)) +
        geom_histogram() + 
        #geom_jitter(alpha = 0.5)
        theme_cowplot() + 
        ggplot.theme


# Number of species not collected in 50 years
data %>% 
    group_by(Species, LastCollected, Specimens) %>%
    summarize() %>% 
    filter(LastCollected < 1970) %>% 
    ungroup() %>%
    summarize(n = n())

names(data)


data %>%
    filter(LastCollected < 1970) %>%
    ggplot() + 
        geom_sf(data = or.counties, color = "black") + 
        geom_point(aes(x = DecimalLongitude, y = DecimalLatitude), 
            color = "darkgreen", size = 2, alpha = 0.6) +
        theme_cowplot() +
        ggplot.theme +
        map.theme


```

### Common Species Temporal Sampling

```{r Common_Species_Herb, echo = F, fig.width=12, fig.height=9}

# 16 most common species in herbarium
data %>% 
    group_by(Family, Species) %>% 
    mutate(Total = n()) %>%
    arrange(desc(Total)) %>% 
    group_by(Family, Species, Total, Decade) %>%
    summarize(Specimens = n()) %>%
    filter(Total > 200) %>%
    ggplot(aes(x = Decade, y = Specimens)) +
        facet_wrap(~ reorder(Species, -Total)) +
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        theme_cowplot() +
        ggtitle("Most Common in OSC Herbarium") + 
        ggplot.theme + 
        theme(strip.text = element_text(face = "bold.italic"))

```


```{r Common_Species_Subset, echo = F, fig.width=10, fig.height=7.5}

# Subset of common/important species
data %>% 
    filter(Species %in% c("Poa secunda", "Eriophyllum lanatum", 
        "Potentilla gracilis", "Quercus garryana", "Acer macrophyllum", 
        "Pseudotsuga menziesii")) %>%
    group_by(Family, Species) %>% 
    mutate(Total = n()) %>% 
    group_by(Family, Species, Total, Decade) %>%
    summarize(Specimens = n()) %>%
    ggplot(aes(x = Decade, y = Specimens)) +
        facet_wrap(~ reorder(Species, -Total)) +
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        theme_cowplot() +
        ggtitle("Common Species") + 
        ggplot.theme + 
        theme(strip.text = element_text(face = "bold.italic"))


```

### Invasive Species Temporal Sampling

```{r Invasives, echo = F, fig.width=10, fig.height=7.5}

data %>% 
    filter(Species %in% c("Rubus armeniacus", "Geranium robertianum", 
        "Hedera helix", "Cytisus scoparius", "Brachypodium sylvaticum", 
        "Cirsium arvense")) %>%
    group_by(Family, Species) %>% 
    mutate(Total = n()) %>% 
    group_by(Family, Species, Total, Decade) %>%
    summarize(Specimens = n()) %>%
    ggplot(aes(x = Decade, y = Specimens)) +
        facet_wrap(~ Species) +
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        theme_cowplot() + 
        ggplot.theme + 
        theme(strip.text = element_text(face = "bold.italic"))

```

## iNaturalist

### Temporal

show increase in sampling

### Spatial

show map

### Invasives


### Common Species


## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```

