---
title: "OSC Herbarium"
author: "James Mickley"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: true
    toc_depth: 3
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```

## Overview

An exploratory analysis of the OSU herbarium's vascular plants.

Primarily, we're concerned with understanding sampling in space, time, and across species. 

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(knitr) # Needed to print tables with kable()
library(taxize) # Needed to derive higher taxonomic ranks with tax_name()
library(sf) # Needed to work with shapefiles

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
data <- read.csv("data/osc-plants.csv")
inat <- read.csv("data/oregon-inat.csv")

# Import Oregon Counties shapefile
or.counties <- read_sf("data/shapefiles/county.shp")


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ ggPlot Theme
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Theme for plots
ggplot.theme <- theme(
    
    # Text size for axis ticks
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    
    # Text size for axis labels
    # Also move them away from the axes a bit for more space
    axis.title.x = element_text(size = 18, face = "bold", vjust = -1),
    axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
    
    # Plot title size, move away from plot, and center
    plot.title = element_text(size = 22, face = "bold", vjust = 4, hjust = 0.5),
    
    # Margins for top, right, bottom, left
    plot.margin = grid::unit(c(1.5, 1.5, 1.5, 1.2), "lines"), 
    
    # Legend text size
    legend.text = element_text(size = 14),
    legend.text.align = 0, 
    legend.title = element_text(size = 16, face = "bold"),
    legend.key.size = grid::unit(1.4, "line"),
    legend.key = element_blank(), 
    
    # Facet label text size and background
    strip.text = element_text(size = 16, face = "bold"),
    strip.background = element_rect(fill = NA)
    )


# Theme for maps to minimize clutter
map.theme <- theme(
    
    # Turn off axes
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
    )


```


```{r Data_Wrangling, echo = F, comment = ""}

# Get higher taxon information from ITIS
families <- data %>% 
    
    # Get a vector of unique family names
    pull(Family) %>% as.character() %>% unique() %>%
    
    # Get the order and division from ITIS
    tax_name(get = c("division", "class", "order"), db = "itis", 
        messages = F) %>% 
    
    # Rename columns
    select(Family = query, Division = division, Class = class, Order = order)


# Clean up the main dataset
data <- data %>%
    
    # Select columns to keep
    select(OccurrenceID, Herbarium, Collection, Catalog, Family, Genus, 
        ScientificName, Genus, SpecificEpithet, Collector, YearCollected, County, 
        DecimalLatitude, DecimalLongitude, CoordinateUncertaintyInMeters) %>% 
    
    # Add a column for decade
    mutate(Decade = floor(YearCollected/10)*10) %>%
    
    # Add a column for species without infraspecific info
    mutate(Species = paste(Genus, SpecificEpithet)) %>%

    # Add higher taxon information
    left_join(families, by = "Family") %>% 
    
    # Filter to vascular plants
    filter(Division == "Tracheophyta") %>%
    
    # Group by species to add some summaries
    group_by(Species) %>%
    
    # Add columns
    mutate(
        
        # Add the number of specimens for each species
        Specimens = n(), 
        
        # Add the last year each species was collected
        LastCollected = first(YearCollected, order_by = desc(YearCollected))
        )


# Clean up the iNaturalist dataset
inat <- inat %>%
  
    # Select columns to keep
    select(observed_on, latitude, longitude, family = taxon_family_name, 
        species = taxon_species_name) %>%
  
    # Make a column for year, derived from collection date
    mutate(Year = format(as.POSIXct(observed_on, format = "%m/%d/%y"), 
        format = "%Y"))


# Clean up the oregon counties shapefile
or.counties <- or.counties %>% 
    
    # Rename NAME column
    select(County = NAME) %>% 
    
    # Transform county names from all caps
    mutate(County = str_to_title(County)) %>% 
    
    # Convert shapefile to wgs84
    st_transform(4326)


# Summarize county data
county.counts <- data %>% 
    
    # Group by species and county
    group_by(Species, County) %>%
    
    # Summarize the number of specimens of each species by county
    summarize(Specimens = n()) %>%
    
    # Group[ by county]
    group_by(County) %>%
    
    # Count the number of species and specimens in each county
    summarize(TotalSpp = n(), Specimens = sum(Specimens)) %>% 
    
    # Arrange counties by number of specimens (most to least)
    arrange(desc(Specimens))

```

## Summaries

### Species

A summary of the number of families, species, and specimens represented, along with stats for the most common families. 

The OSC herbarium has just over 4500 species of vascular plant in Oregon with good data.

```{r State_Summary, echo = F, comment = ""}

# Summarize total families, species, and specimens in dataset
data %>% 
    
    # Group by species
    group_by(Family, Species) %>%
    
    # Count up the number of specimens for each species
    summarize(Specimens = n()) %>%

    # Group by family
    group_by(Family) %>%
        
    # Summarize the # of species and specimens by family
    summarize(Species = n(), Specimens = sum(Specimens)) %>%
        
    # Summarize the # of families
    summarize(Families = n(), Species = sum(Species), 
        Specimens = sum(Specimens)) %>%

    # Output a table of the data
    kable(format = "markdown")  


# Summarize for common families (> 2000 specimens)
data %>% 
    
    # Group by species
    group_by(Family, Species) %>%
    
    # Count up the number of specimens for each species
    summarize(Specimens = n()) %>%

    # Group by family
    group_by(Family) %>%
        
    # Summarize the # of species and specimens by family
    summarize(Species = n(), Specimens = sum(Specimens)) %>%
    
    # Arrange families by the # of specimens descending
    arrange(desc(Specimens)) %>%
    
    # Show the top 20 families
    head(20) %>%

    # Output a table of the data
    kable(format = "markdown")  


```

### Locations

A map of all the specimen records to look at sampling intensity across space. 

It looks like there's less sampling in the mountains, unsurprisingly. 

```{r State_Map, echo = F, comment = ""}

# Map all specimen records
data %>%
    ggplot() + 
        
        # Add the oregon counties outline
        geom_sf(data = or.counties, color = "black") + 
  
        # Add points for each specimen location
        geom_point(aes(x = DecimalLongitude, y = DecimalLatitude), 
            color = "darkgreen", size = 2, alpha = 0.3) +
  
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme +
  
        # Add map theme
        map.theme

```

### Collectors

A list of important collectors for Oregon specimens in OSC, with active dates.

Morton Peck of WILLU tops the list (Manual of the Higher Plants of OR). Louis Henderson of UofO is second, and Richard Halse is third.

Some great information on the background of these collectors is [here](https://bpp.oregonstate.edu/herbarium/databases/collectors-specimen-database).

```{r Collectors, echo = F, comment = ""}

# Parse out the important collectors
data %>%
    
    # Count the number of specimens, first and last year for each collector
    group_by(Collector) %>%
    mutate(Specimens = n(), FirstAccession = min(YearCollected), 
        LastAccession = max(YearCollected)) %>%
    
    # Group by collector & species, summarize to one row per species-collector
    group_by(Collector, Species, Specimens, FirstAccession, 
        LastAccession) %>%
    summarize() %>%
    
    # Group by collector and summarize the number of species
    group_by(Collector, Specimens, FirstAccession, LastAccession) %>%
    summarize(NumSpecies = n()) %>%
    
    # Reorder columbns
    select(Collector, Specimens, NumSpecies, FirstAccession, LastAccession) %>%
    
    # Arrange by the number of specimens descending
    arrange(desc(Specimens)) %>%
    
    # Pick out the top 20 collectors
    head(20) %>%
    
    # Output a table of the data
    kable(format = "markdown")  

```


## Sampling Biases

Let's look at spatial biases in sampling in three axes: spatial, temporal, and species

### Spatial

Counties vary a lot in the number of specimens, over 9000 to less than 500. This is pretty typical for an herbarium. 

Lane and Benton Counties are well represented, so are Harney, Wallowa, and Josephine.

```{r Specimens_County, echo = F}

# Bar graph of specimens by county
county.counts %>%
    
    # Plot specimens against county ordered by # of specimens
    ggplot(aes(x = reorder(County, -Specimens), y = Specimens)) + 
    
        # Add bar graph
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
          alpha = 0.2) + 
    
        # Add x-axis label
        labs(x = "County") +
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme +
  
        theme(axis.text.x = element_text(angle = 90, vjust = 0.2))


# Add the number of specimens per county to the county shapefile
or.counties %>% left_join(county.counts) %>%
    
    # Make a map of the number of specimens by county
    ggplot() + 
    
        # Color the counties by the number of specimens
        geom_sf(aes(fill = Specimens), color = "black") + 
    
        # Add county name labels
        geom_sf_label(aes(label = County), size = 3, fontface = "bold") + 
    
        # Set the color scale
        scale_fill_gradient(low = "darkseagreen1", high = "darkgreen") +

        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme +
  
        # Add map theme
        map.theme


        
```

Species diversity at the county level looks partly tied to the number of specimens. Counties with more specimens have more diversity. This suggests that sampling is incomplete at the county level; we'd like to see diversity be independent of sampling.

```{r Diversity_County, echo = F}

# Add the number of specimens per county to the county shapefile
or.counties %>% left_join(county.counts) %>%
    
    # Make a map of the number of species by county
    ggplot() + 
        
        # Color the counties by the number of specimens
        geom_sf(aes(fill = TotalSpp), color = "black") + 
    
        # Add county name labels
        geom_sf_label(aes(label = County), size = 3, fontface = "bold") + 
    
        # Set the color scale
        scale_fill_gradient(low = "darkseagreen1", high = "darkgreen") +
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme +
  
        # Add map theme
        map.theme


```

### Temporal

Show a fine-grained analysis of the number of specimens collected each year. There's a lot of variation, but there's been a reduction in the last five years or so.

```{r Specimens_Year, echo = F}

# Show specimens by year
data %>% 
    
    # Group by year and count up the number of specimens per year
    group_by(YearCollected) %>% 
    summarize(Specimens = n()) %>%
    
    # Plot the number of specimens by year collected
    ggplot(aes(x = YearCollected, y = Specimens)) + 
    
        # Add a bar plot
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme


```

Simplifying a bit, we can look at specimens by decade. The best sampling was 1910-1950. The 2010s is the lowest decade since the 1890s.


```{r Specimens_Decade, echo = F}

# Show specimens by decade
data %>% 
    
    # Group by decade and count up the number of specimens per decade
    group_by(Decade) %>% 
    summarize(Specimens = n()) %>%
    
    # Plot the number of specimens by decade
    ggplot(aes(x = Decade, y = Specimens)) + 
    
        # Add bar plot
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme



```

### Spatiotemporal

We can also get a sense of specimens by decade by county to see if there are any spatiotemporal trends. For the most part, counties follow the broader temporal pattern. Multnomah, Klamath, and Marion have hardly been sampled in a long time.

```{r Specimens_County_Decade, echo = F, fig.width = 12, fig.height = 9}

# Show specimens by decade by county
data %>% 
    
    # Group by decade and county, summarize the number of specimens per decade
    group_by(County, Decade) %>% 
    summarize(Specimens = n()) %>%
    
    # Plot the number of specimens per decade
    ggplot(aes(x = Decade, y = Specimens)) + 
    
        # Split up graph by county
        facet_wrap(~ County) +
    
        # Add barplot
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
        
        # Limit the number of x-axis ticks
        scale_x_continuous(breaks = c(1900,1950, 2000)) +
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme


```

Focus on the six most-sampled counties

```{r Specimens_County_Decade_Subset, echo = F}

# Show specimens by decade for most common counties
data %>% 
    
    # Filter to the six most common counties
    filter(County %in% c("Lane", "Harney", "Wallowa", "Josephine", 
        "Benton", "Jackson")) %>% 

    # Group by decade and county, summarize the number of specimens per decade
    group_by(County, Decade) %>% 
    summarize(Specimens = n()) %>%
    
    # Plot the number of specimens per decade
    ggplot(aes(x = Decade, y = Specimens)) + 
    
        # Split up graph by county
        facet_wrap(~ reorder(County, -Specimens)) +
    
        # Add barplot
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme

```

### Undersampled Species

Undersampled Species
One way to get at data-deficient species, is to look at the number of species with very few records.

Of the 4500 species recorded, over 2000 less than 10 specimens statewide, and nearly 1000 species are only represented by 1-2 specimens. These would be a good target, though it's possible some are outdated names or something of that sort.


```{r Undersampled_Species, echo = F}

# Filter to only include species with 10 or fewer specimens
data %>% 
    
    # Summarize number of specimens per species
    group_by(Species, Specimens) %>%
    summarize() %>%
    
    filter(Specimens <= 10) %>%
    
    # Group by the number of specimens
    group_by(Specimens) %>% 
    
    # Summarize the number of species with a certain number of specimens
    summarize(Species = n()) %>% 
    
    # Output a table of the data
    kable(format = "markdown")



# Graph the number of specimens per species
data %>% 
    
    # Summarize number of specimens per species
    group_by(Species) %>%
    summarize(Specimens = n()) %>%

    # Calculate the median specimens per species
    mutate(MedianSp = median(Specimens)) %>%
  
    # Plot histogram of specimens per species
    ggplot(aes(x = Specimens)) + 
        
        # Add histogram  
        geom_histogram(bins = 30, fill = "darkgreen", color = "darkgreen", 
            alpha = 0.2) +
    
        # Add median specimens per species
        geom_vline(aes(xintercept = MedianSp), color = "red", size = 1.2) +
        
        # Add axis labels
        labs(x = "Number of Specimens", y = "Number of Species") +

        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme


```

Another way that we can look at sampling across species is to add a time component and look for species that have not been collected in a long time. 

When we look at when each species was last collected, the median is 1996, so half of the species have not been collected in about 25 years. There's a significant proportion that have not been collected in over 50 years too (1970).

Again, these could be old names, but they could also be species that have become rare or extinct, or simply haven't been looked for, because of inaccessibility or lack of taxonomic expertise.

```{r Undersampled_Species_LastCollected, echo = F}

data %>% 
    
    # Group by species, and summarize
    group_by(Species, LastCollected, Specimens) %>%
    summarize() %>% 
    
    # Ungroup
    ungroup() %>%
    
    # Limit to species with 10 or fewer herbarium specimens
    filter(Specimens <= 10) %>%
    
    # Calculate the median last collected year
    mutate(MedianYr = median(LastCollected)) %>%
    
    # Plot the last collected date
    ggplot(aes(x = LastCollected)) +
    
        # Add histogram
        geom_histogram(bins = 30, color = "darkgreen", fill = "darkgreen", 
              alpha = 0.2) + 

        # Add labels
        labs(x = "Last Year Collected", y = "Number of Species") +
    
        # Add median last collected date
        geom_vline(aes(xintercept = MedianYr), color = "red", size = 1.2) +
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme


```



A dataset of species not collected in 50 years species, with the last year they were collected is available at [oldcollections.csv](/data/oldcollections.csv)

The actual number of species not collected in 50 years:

```{r LastCollected_Count, echo = F, comment = ""}

# Compile a list of species that have not been collected in 50 years
data %>% 
    
    # Group by Species and summarize
    group_by(Family, Species, LastCollected, Specimens) %>%
    summarize() %>% 
    
    # Filter to only those species that haven't been collected in 50 years
    filter(LastCollected < 1970) %>% 
    
    # Ungroup
    ungroup() %>%
    
    # Sort by family and then by species alphabetically
    arrange(Family, Species) %>%

    # Write the list of species to CSV
    write.csv("data/oldcollections.csv", row.names = F)


# Count the number of species not collected in 50 years
data %>% 
    
    # Group by Species and summarize
    group_by(Species, LastCollected, Specimens) %>%
    summarize() %>% 
    
    # Filter to only those species that haven't been collected in 50 years
    filter(LastCollected < 1970) %>% 
    
    # Ungroup
    ungroup() %>%
    
    # Count the number of species
    summarize(Species_Not_Collected_In_50_Years = n()) %>%
    
    # Output a table of the data
    kable(format = "markdown")

```

Here's a map of where species that have not been collected in 50 years were originally collected.

```{r LastCollected_Map, echo = F}

# Show map of species not collected in 50 years
data %>%
    
    # Filter to species not collected in 50 years
    filter(LastCollected < 1970) %>%
    
    # Plot map of where these species were collected
    ggplot() + 
    
        # Add oregon counties outline
        geom_sf(data = or.counties, color = "black") + 
    
        # Add points for specimens of these species
        geom_point(aes(x = DecimalLongitude, y = DecimalLatitude), 
            color = "darkgreen", size = 2, alpha = 0.6) +
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme +
  
        # Add map theme
        map.theme


```

### Common Species Temporal Sampling

This shows temporal sampling for the 16 most common species in the herbarium

```{r Common_Species_Herb, echo = F, fig.width=12, fig.height=9}

# 16 most common species in herbarium
data %>% 
    
    # Group by species and find the total specimens per species
    group_by(Species) %>% 
    mutate(Total = n()) %>% 
    
    # Group by species and decade and count the number of specimens/species
    group_by(Species, Total, Decade) %>%
    summarize(Specimens = n()) %>%
    
    # Arrange by total specimens per species descending
    arrange(desc(Total)) %>% 
    
    # Only keep species with more than 200 specimens
    filter(Total > 200) %>%
    
    # Plot specimes by decade
    ggplot(aes(x = Decade, y = Specimens)) +
        
        # Facet by our target species
        facet_wrap(~ reorder(Species, -Total)) +
    
        # Add bar graph
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 

        # Add graph title
        labs(title = "Most Common in OSC Herbarium") + 
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme + 

        # Make species names italic in facet labels
        theme(strip.text = element_text(face = "bold.italic"))

```

We can focus in on the three most common species. Also, I've included three tree species that are relatively common in the landscape for comparison.

```{r Common_Species_Subset, echo = F, fig.width=10, fig.height=7.5}

# Subset of common/important species
data %>% 
    
    # Filter to 1) three most common species and 2) important trees
    filter(Species %in% c("Poa secunda", "Eriophyllum lanatum", 
        "Potentilla gracilis", "Quercus garryana", "Acer macrophyllum", 
        "Pseudotsuga menziesii")) %>%
    
    # Group by species and find the total specimens per species
    group_by(Species) %>% 
    mutate(Total = n()) %>% 
    
    # Group by species and decade and count the number of specimens/species
    group_by(Species, Total, Decade) %>%
    summarize(Specimens = n()) %>%
    
    # Plot specimens by decade
    ggplot(aes(x = Decade, y = Specimens)) +
    
        # Facet by our target species
        facet_wrap(~ reorder(Species, -Total)) +
    
        # Add bar graph
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 

        # Add graph title
        labs(title = "Common Species") +
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme + 

        # Make species names italic in facet labels
        theme(strip.text = element_text(face = "bold.italic"))


```

### Invasive Species Temporal Sampling

Likewise, we can look at temporal sampling for some common invasives. 
These invasives are not well sampled at all, regardless of decade. They're another possible focal point of targeted collecting.


```{r Invasives, echo = F, fig.width=10, fig.height=7.5}

# Plot of invasive species by decade
data %>% 
    
    # Filter to include six invasive species
    filter(Species %in% c("Rubus armeniacus", "Geranium robertianum", 
        "Hedera helix", "Cytisus scoparius", "Brachypodium sylvaticum", 
        "Cirsium arvense")) %>%
    
    # Group by species and find the total specimens per species
    group_by(Species) %>% 
    mutate(Total = n()) %>% 
    
    # Group by species and decade and count the number of specimens/species
    group_by(Species, Total, Decade) %>%
    summarize(Specimens = n()) %>%

    # Plot specimens by decade
    ggplot(aes(x = Decade, y = Specimens)) +
    
        # Facet by our target species
        facet_wrap(~ Species) +
    
        # Add bar graph
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme + 

        # Make species names italic in facet labels
        theme(strip.text = element_text(face = "bold.italic"))

```

## iNaturalist

By comparison to the herbarium data, iNaturalist observations are growing exponentially. 

iNaturalist has collected more observations in Oregon in each of the last four years than the best year for the herbarium, and last year it was 60,000!

```{r iNat_Year, echo = F}

# Graph iNaturalist data by year
inat %>% 
    
    # Group observations by year and count the number for each year
    group_by(Year) %>%
    summarize(Observations = n()) %>%
    
    # Filter to only the last 10 years
    filter(Year > 2009, Year < 2020) %>%
    
    # Plot # of observations against year
    ggplot(aes(x = Year, y = Observations)) + 
    
        # Add bar graph
        geom_bar(stat = "identity", color = "darkgreen", fill = "darkgreen", 
            alpha = 0.2) + 
    
        # Add cowplot minimal theme
        theme_cowplot() +
  
        # Ad custom theme
        ggplot.theme


```


## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```

